---
title: 'Growth Curve Analysis'
author: 'Angie Vreugdenhil-Hayslette, MSc.'
date: '2/17/2020'
output: html_document
---

REFRESH

```{r start, include=FALSE}
# refresh
rm(list=ls())
# setup Knit
knitr::opts_chunk$set(echo=TRUE)
```

LOAD PACKAGES

```{r packages, include=FALSE, echo=FALSE}
# load required packages
library(plyr)
library(tidyr) 
library(dplyr)
library(tibble)
library(viridis)
library(ggplot2)
library(tinytex)
library(latexpdf)
library(data.table)
library(growthrates)
```

SETUP

Actions:
Change the project name. 
Choose wells/timepoints that may need to be removed. 
Choose the wells with your blanks.

```{r input, include=FALSE, echo=FALSE}
# project name
project<- 'miniterm'
# project folder
folder <- '/cloud/project/miniterm/'
# folder with meta
foldermeta <- '/cloud/project/miniterm/meta/'
# metafile name
meta <- 'meta_example_BioScreen.csv'
# folder with data
folderdata <- '/cloud/project/miniterm/data/'
# datafile name
data <- 'data_example_BioScreen.csv'
# wells to be removed based on observation (do not remove = 1000)
well1<- 1000
well2<- 1000
well3<- 1000
# timepoints to be removed based on observation (do not remove = 1000)
time1<- 1000
time2<- 1000
time3<- 1000
# blank wells (not a blank = 1000)
blank1<- 1000
blank2<- 1000
blank3<- 1000
blank4<- 1000
blank5<- 1000
blank6<- 1000
blank7<- 1000
blank8<- 1000
blank9<- 1000
blank10<- 1000
blank11<- 1000
blank12<- 1000
```

DATA INPUT - BioScreen

Actions:
If amount of wells/timepoint to be removed has been altered, make sure to include all below.
If amount of blanks has been altered, make sure to include all below.
Make sure the timeframe is set correctly.
Set the subtracted value of the mean blank if lowest mean value is negative.

```{r load, include=FALSE, echo=FALSE}
# load data (mt0 and dt0)
## read in metafile
mt <- read.csv(paste(foldermeta, meta, sep = ''),
               strip.white=TRUE,
               sep = ',',
               header = TRUE)
mt0 <- mt[-1, -(1:2)]
## read in datafile
dt <- read.csv(paste(folderdata, data, sep = ''),
               strip.white=TRUE,
               sep = ',',
               header = FALSE)
dt0 <- dt[-(1:2), -2]
colnames(dt0) <- c('time', dt0[1, -1])
dt0 <- dt0[-1, ]

# edit data (c_mtdt)
## remove wells based on observation in meta
mt0 <- mt0[!(mt0$variable == well1 |
               mt0$variable == well2 |
               mt0$variable == well3),]
## remove wells based on observation in data
dt0 <- dt0 %>% select(-contains(as.character(well1)))
dt0 <- dt0 %>% select(-contains(as.character(well2)))
dt0 <- dt0 %>% select(-contains(as.character(well3)))
## extract hours, mins, secs
hou <-
  as.numeric(lapply(strsplit(as.character(dt0$time), ':'), '[[', 1))
min <-
  as.numeric(lapply(strsplit(as.character(dt0$time), ':'), '[[', 2))
sec <-
  as.numeric(lapply(strsplit(as.character(dt0$time), ':'), '[[', 3))
## set total time to hours
dt0$time <- (hou) + min / 60 + (sec / 3600)
## remove timepoints based on observation in data
dt0<-dt0[!(dt0$time==time1),]
dt0<-dt0[!(dt0$time==time2),]
dt0<-dt0[!(dt0$time==time3),]
## melt and merge the dataframe to one list
m_dt <- melt(dt0, id = 'time')
mtdt <- merge(m_dt, mt0, by = 'variable')
## fix delta symbol by replacing default '_' with delta symbol
mtdt$name<- gsub('_','\u0394',mtdt$name)
## remove and rename columns in dataframe
mtdt[is.na(mtdt)]<- ''
mtdt$environment <- ifelse(mtdt$condition=='',as.character(mtdt$media),paste(mtdt$media,'+',mtdt$condition,sep=''))
mtdt$full <- ifelse(mtdt$name=='',as.character(mtdt$strain),paste(as.character(mtdt$strain),' ',mtdt$name,sep=''))
mtdt$ID <- ifelse(mtdt$name=='' & mtdt$condition=='', as.character(mtdt$strain),
                  ifelse(mtdt$condition=='',paste(as.character(mtdt$strain),' ',mtdt$name,sep=''),
                         ifelse(mtdt$name=='',paste(as.character(mtdt$strain),' (',as.character(mtdt$condition),')',sep=''),
                                paste(as.character(mtdt$strain),' ',mtdt$name,' (',as.character(mtdt$condition),')',sep=''))))
well <- paste(mtdt$variable, mtdt$strain, sep = ' ')
mtdt <- cbind(well, mtdt)
### remove rows without ID
c_mtdt <- mtdt[!(mtdt$ID == ''), ]

# edit timepoints (t_mtdt)
## display min and max timepoints
min(c_mtdt$time)
max(c_mtdt$time)
## set time boundaries to be included in all analysis
t_mtdt <- c_mtdt[!(c_mtdt$time <= 3 | c_mtdt$time >= 48), ]

# select blanks (blank_all)
blanks<- t_mtdt[(t_mtdt$variable == blank1 |
                   t_mtdt$variable == blank2 |
                   t_mtdt$variable == blank3 |
                   t_mtdt$variable == blank4 |
                   t_mtdt$variable == blank5 |
                   t_mtdt$variable == blank6), ]
## set baseline
blank_all<- ddply(blanks, c('time','environment'),
                  summarize,
                  b_mn=mean(value)) #subtracted value prevents negative values
## minimum blank value
min(blank_all$b_mn)
## maximum blank value
max(blank_all$b_mn)

# select data WITHOUT blank correction (tm_mtdt)
## remove wells that aren't sample nor a blank
tc_mtdt<- t_mtdt[!t_mtdt$name=='blank',]
tu_mtdt<- rbind(tc_mtdt,blanks)
## summarize all values (time, strain, media, condition, biorep)
tm_mtdt<- ddply(tu_mtdt, c('time', 'ID', 'full', 'strain', 'name', 'media', 'condition', 'environment', 'biorep'),
                summarize,
                value_mn = mean(value),
                value_sd = sd(value))

# select data WITH blank correction (b_mtdt)
## add correct blank values to each row 
b_mtdt<- merge(tm_mtdt,blank_all,c('time','environment'))
head(b_mtdt)
## reorganize columns (variable number subtracted to prevent negative blank values)
b_mtdt$value_mn<- b_mtdt$value_mn - (b_mtdt$b_mn - 0.005)
b_mtdt$b_mn<- NULL
## remove blank rows
b_mtdt <- b_mtdt[!b_mtdt$name=='blank',]
## if this value is negative, subtract it from the mean above
min(b_mtdt$value_mn)
```

PLOT BLANKS

```{r blank, include=FALSE, echo=TRUE}
# folder for plots
 dir.create(paste(folder,'plots',sep=''))
 setwd(paste(folder,'plots',sep=''))

# blanks
 p_blanks<- 
  ggplot(blanks, aes(x = time, y = value, color = environment)) +
  ggtitle('Blank curves') +
  xlab('time (h)') + 
  ylab(bquote(bold(paste('Log'['2'],'(Absorbance at 600nm)')))) +
  geom_line() +
  scale_y_continuous(trans = 'log2') +
  scale_color_viridis('condition', discrete = TRUE) +
  theme(aspect.ratio=1) +
  theme(#legend.position = 'none',
    strip.text = element_text(face = "italic",size = 8),
    plot.title = element_text(face='bold', color='#000000', size=12, angle=0),
    axis.title.x = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.title.y = element_text(face='bold', color='#000000', size=10, angle=90),
    legend.title = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.text.y = element_text(face='plain', color='#000000', size=8, angle=0),
    axis.text.x = element_text(face='plain', color='#000000', size=8, angle=0),
    legend.text = element_text(face='plain', color='#000000', size=10, angle=0))

ggsave(p_blanks,
       file = paste(
         format(Sys.time(), '%Y%m%d%H%M'),
         '_',
         project,
         '_',
         'plot_blanks',
         '.jpeg',
         sep = ''),
       width = 10, height = 8, units = "in", device='jpeg')
```

PLOT RAW - grid of all wells

```{r grid, include=FALSE, echo=TRUE}
# folder for plots
 dir.create(paste(folder,'plots',sep=''))
 setwd(paste(folder,'plots',sep=''))

# growthcurves - grid
p_gri <-
  ggplot(tu_mtdt, aes(x = time, y = value)) +
  ggtitle('Growthcurves each well') +
  xlab('time (h)') + 
  ylab(bquote(bold(paste('Log'['2'],'(Absorbance at 600nm)')))) +
  geom_line() +
  scale_y_continuous(trans = 'log2') +
  facet_wrap(~ well) +
  theme(#legend.position = 'none',
    strip.text = element_text(face = "plain",size = 8),
    plot.title = element_text(face='bold', color='#000000', size=12, angle=0),
    axis.title.x = element_text(face='bold', color='#000000', size=8, angle=0),
    axis.title.y = element_text(face='bold', color='#000000', size=8, angle=90),
    legend.title = element_text(face='bold', color='#000000', size=8, angle=0),
    axis.text.y = element_text(face='plain', color='#000000', size=6, angle=0),
    axis.text.x = element_text(face='plain', color='#000000', size=6, angle=0),
    legend.text = element_text(face='plain', color='#000000', size=8, angle=0))

ggsave(p_gri, file = paste(
  format(Sys.time(), '%Y%m%d%H%M'),
  '_',
  project,
  '_',
  'plot_raw_grid_all',
  '.jpeg',
  sep = ''),
  width = 10, height = 8, units = "in", device='jpeg')
```

PLOT RAW DATA - every well in one plot

```{r raw, include=FALSE, echo=TRUE}
# folder for plots
dir.create(paste(folder,'plots',sep=''))
setwd(paste(folder,'plots',sep=''))

# growthcurves - one plot - well
p_well <-
  ggplot(tu_mtdt, aes(x = time, y = value, color = well)) +
  ggtitle('Growthcurves each well') +
  xlab('time (h)') + 
  ylab(bquote(bold(paste('Log'['2'],'(Absorbance at 600nm)')))) +
  geom_line() +
  scale_y_continuous(trans = 'log2') +
  scale_color_viridis('well', discrete = TRUE) +
  theme(aspect.ratio=1) +
  theme(legend.position = 'none',
        strip.text = element_text(face = "italic",size = 8),
        plot.title = element_text(face='bold', color='#000000', size=12, angle=0),
        axis.title.x = element_text(face='bold', color='#000000', size=10, angle=0),
        axis.title.y = element_text(face='bold', color='#000000', size=10, angle=90),
        legend.title = element_text(face='bold', color='#000000', size=10, angle=0),
        axis.text.y = element_text(face='plain', color='#000000', size=8, angle=0),
        axis.text.x = element_text(face='plain', color='#000000', size=8, angle=0),
        legend.text = element_text(face='plain', color='#000000', size=10, angle=0))

ggsave(p_well, file = paste(
  format(Sys.time(), '%Y%m%d%H%M'),
  '_',
  project,
  '_',
  'plot_raw_well_all',
  '.jpeg',
  sep = ''),
  width = 10, height = 8, units = "in", device='jpeg')
```

PLOT RAW DATA - averaged on ID

```{r id, include=FALSE, echo=TRUE}
# folder for plots
dir.create(paste(folder,'plots',sep=''))
setwd(paste(folder,'plots',sep=''))

# growthcurves - one plot - ID
p_id <-
  ggplot(tu_mtdt, aes(x = time, y = value, color = ID)) +
  ggtitle('Growthcurves each strain') +
  xlab('time (h)') + 
  ylab(bquote(bold(paste('Log'['2'],'(Absorbance at 600nm)')))) +
  geom_point() +
  scale_y_continuous(trans = 'log2') +
  scale_color_viridis('strain description', discrete = TRUE) +
  theme(aspect.ratio=1) +
  theme(#legend.position = 'none',
    strip.text = element_text(face = "italic",size = 8),
    plot.title = element_text(face='bold', color='#000000', size=12, angle=0),
    axis.title.x = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.title.y = element_text(face='bold', color='#000000', size=10, angle=90),
    legend.title = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.text.y = element_text(face='plain', color='#000000', size=8, angle=0),
    axis.text.x = element_text(face='plain', color='#000000', size=8, angle=0),
    legend.text = element_text(face='italic', color='#000000', size=10, angle=0))

ggsave(p_id, file = paste(
  format(Sys.time(), '%Y%m%d%H%M'),
  '_',
  project,
  '_',
  'plot_raw_ID_all',
  '.jpeg',
  sep = ''),
  width = 10, height = 8, units = "in", device='jpeg')
```

PLOT RAW DATA - averaged on ID (blank corrected)

```{r idbc, include=FALSE, echo=TRUE}
# folder for plots
dir.create(paste(folder,'plots',sep=''))
setwd(paste(folder,'plots',sep=''))

# # growthcurves - one plot - ID
p_b_id <-
  ggplot(b_mtdt, aes(x = time, y = value_mn, color = ID)) +
  ggtitle('Growthcurves each strain') +
  xlab('time (h)') + 
  ylab(bquote(bold(paste('Log'['2'],'(Absorbance at 600nm)')))) +
  geom_point() +
  scale_y_continuous(trans = 'log2') +
  scale_color_viridis('strain description', discrete = TRUE) +
  theme(aspect.ratio=1) +
  theme(#legend.position = 'none',
    strip.text = element_text(face = "italic",size = 8),
    plot.title = element_text(face='bold', color='#000000', size=12, angle=0),
    axis.title.x = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.title.y = element_text(face='bold', color='#000000', size=10, angle=90),
    legend.title = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.text.y = element_text(face='plain', color='#000000', size=8, angle=0),
    axis.text.x = element_text(face='plain', color='#000000', size=8, angle=0),
    legend.text = element_text(face='italic', color='#000000', size=10, angle=0))

ggsave(p_b_id, file = paste(
  format(Sys.time(), '%Y%m%d%H%M'),
  '_',
  project,
  '_',
  'plot_bcor_ID_all',
  '.jpeg',
  sep = ''),
  width = 10, height = 8, units = "in", device='jpeg')
```

PLOT CONFIDENCE INTERVAL

Here we calculate the 95% confidence interval. This percentage can be change to any other by adjusting the conf_int function.

```{r stats, include=FALSE, echo=TRUE}
# folder for plots
dir.create(paste(folder,'plots',sep=''))
setwd(paste(folder,'plots',sep=''))

# confidence interval function: set to 95%
conf_int <- function(data) {
  n <- length(data)
  error <- qt(0.975, df = n - 1) * sd(data) / sqrt(n)
  return(error)
}

stats <- tm_mtdt %>%
  group_by(full, ID, strain, name, environment, time) %>%
  summarise(
    reps = length(value_mn),
    average = mean(value_mn),
    CI = conf_int(value_mn)) %>%
  filter(!strain == '') #remove blanks

# growth curves - average strain
p_ave_str <-
  ggplot(stats, aes(x = time, y = average, color = environment)) +
  ggtitle('Growthcurves averaged each condition and seperated in strain') +
  xlab('time (h)') + 
  ylab(bquote(bold(paste('Log'['2'],'(Absorbance at 600nm)')))) +
  geom_ribbon(
    aes(
      ymin = average - CI,
      ymax = average + CI,
      fill = environment), color = NA, alpha = 0.3) +
  geom_line() +
  scale_y_continuous(trans = 'log2') +
  facet_grid(. ~ full) +
  scale_fill_viridis('condition', discrete = TRUE) +
  scale_color_viridis('condition', discrete = TRUE) +
  theme(aspect.ratio=1) +
  theme(#legend.position = 'none',
    strip.text = element_text(face = "italic",size = 7),
    plot.title = element_text(face='bold', color='#000000', size=12, angle=0),
    axis.title.x = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.title.y = element_text(face='bold', color='#000000', size=10, angle=90),
    legend.title = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.text.y = element_text(face='plain', color='#000000', size=8, angle=0),
    axis.text.x = element_text(face='plain', color='#000000', size=8, angle=0),
    legend.text = element_text(face='plain', color='#000000', size=10, angle=0))

ggsave(p_ave_str,
       file = paste(
         format(Sys.time(), '%Y%m%d%H%M'),
         '_',
         project,
         '_',
         'plot_ci_strain',
         '.jpeg',
         sep = ''),
       width = 10, height = 8, units = "in", device='jpeg')

# growth curves - average condition
p_ave_con <-
  ggplot(stats, aes(x = time, y = average, color = full)) +
  ggtitle('Growthcurves averaged each strain and seperated in condition') +
  xlab('time (h)') + 
  ylab(bquote(bold(paste('Log'['2'],'(Absorbance at 600nm)')))) +
  geom_ribbon(
    aes(
      ymin = average - CI,
      ymax = average + CI,
      fill = full), color = NA, alpha = 0.3) +
  geom_line() +
  scale_y_continuous(trans = 'log2') +
  facet_grid(. ~ environment) +
  scale_fill_viridis('Strain name', discrete = TRUE) +
  scale_color_viridis('Strain name', discrete = TRUE) +
  theme(aspect.ratio=1) +
  theme(#legend.position = 'none',
    strip.text = element_text(face = "italic",size = 10),
    plot.title = element_text(face='bold', color='#000000', size=12, angle=0),
    axis.title.x = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.title.y = element_text(face='bold', color='#000000', size=10, angle=90),
    legend.title = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.text.y = element_text(face='plain', color='#000000', size=8, angle=0),
    axis.text.x = element_text(face='plain', color='#000000', size=8, angle=0),
    legend.text = element_text(face='italic', color='#000000', size=10, angle=0))

ggsave(p_ave_con,
       file = paste(
         format(Sys.time(), '%Y%m%d%H%M'),
         '_',
         project,
         '_',
         'plot_ci_condition',
         '.jpeg',
         sep = ''),
       width = 10, height = 8, units = "in", device='jpeg')
```

PLOT CONFIDENCE INTERVAL (blank corrected)

Here we calculate the 95% confidence interval. This percentage can be change to any other by adjusting the conf_int function.

```{r statsbc, include=FALSE, echo=TRUE}
# folder for plots
dir.create(paste(folder,'plots',sep=''))
setwd(paste(folder,'plots',sep=''))

# confidence interval function: set to 95%
conf_int <- function(data) {
  n <- length(data)
  error <- qt(0.975, df = n - 1) * sd(data) / sqrt(n)
  return(error)
}

stats <- b_mtdt %>%
  group_by(full, ID, strain, name, environment, time) %>%
  summarise(
    reps = length(value_mn),
    average = mean(value_mn),
    CI95 = conf_int(value_mn)) %>%
  filter(!strain == '')

# growth curves - average strain
pb_ave_str_b <-
  ggplot(stats, aes(x = time, y = average, color = environment)) +
  ggtitle('Growthcurves averaged each condition and seperated by strain') +
  xlab('time (h)') + 
  ylab(bquote(bold(paste('Log'['2'],'(Absorbance at 600nm)')))) +
  geom_ribbon(
    aes(
      ymin = average - CI,
      ymax = average + CI,
      fill = environment), color = NA, alpha = 0.3) +
  geom_line() +
  scale_y_continuous(trans = 'log2') +
  facet_grid(. ~ full) +
  scale_fill_viridis('condition', discrete = TRUE) +
  scale_color_viridis('condition', discrete = TRUE) +
  theme(aspect.ratio=1) +
  theme(#legend.position = 'none',
    strip.text = element_text(face = "italic",size = 7),
    plot.title = element_text(face='bold', color='#000000', size=12, angle=0),
    axis.title.x = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.title.y = element_text(face='bold', color='#000000', size=10, angle=90),
    legend.title = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.text.y = element_text(face='plain', color='#000000', size=8, angle=0),
    axis.text.x = element_text(face='plain', color='#000000', size=8, angle=0),
    legend.text = element_text(face='plain', color='#000000', size=10, angle=0))

ggsave(pb_ave_str_b,
       file = paste(
         format(Sys.time(), '%Y%m%d%H%M'),
         '_',
         project,
         '_',
         'plot_ci_strain_bcor',
         '.jpeg',
         sep = ''),
       width = 10, height = 8, units = "in", device='jpeg')

# growth curves - average condition
p_ave_con_b <-
  ggplot(stats, aes(x = time, y = average, color = full)) +
  ggtitle('Growthcurves averaged each strain and seperated by condition') +
  xlab('time (h)') + 
  ylab(bquote(bold(paste('Log'['2'],'(Absorbance at 600nm)')))) +
  geom_ribbon(
    aes(
      ymin = average - CI,
      ymax = average + CI,
      fill = full), color = NA, alpha = 0.3) +
  geom_line() +
  scale_y_continuous(trans = 'log2') +
  facet_grid(. ~ environment) +
  scale_fill_viridis('strain', discrete = TRUE) +
  scale_color_viridis('strain', discrete = TRUE) +
  theme(aspect.ratio=1) +
  theme(#legend.position = 'none',
    strip.text = element_text(face = "italic",size = 8),
    plot.title = element_text(face='bold', color='#000000', size=12, angle=0),
    axis.title.x = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.title.y = element_text(face='bold', color='#000000', size=10, angle=90),
    legend.title = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.text.y = element_text(face='plain', color='#000000', size=8, angle=0),
    axis.text.x = element_text(face='plain', color='#000000', size=8, angle=0),
    legend.text = element_text(face='italic', color='#000000', size=10, angle=0))

ggsave(p_ave_con_b,
       file = paste(
         format(Sys.time(), '%Y%m%d%H%M'),
         '_',
         project,
         '_',
         'plot_ci_condition_bcor',
         '.jpeg',
         sep = ''),
       width = 10, height = 8, units = "in", device='jpeg')
```

PLOT MUMAX

For this plot we will first model the growth using a algorithm of Kahm et al. 2010 (see resources folder). The spline fitting is done with log-transformed data, assuming exponential growth at the time point of the maximum of its first derivative. The maximum growthrate is estimated based on this fit.

Actions:
Make sure to change the par setting in the first plot based on the amount of plots included.

```{r mumax, include=FALSE, echo=TRUE}
# folder for plots
dir.create(paste(folder,'plots',sep=''))
setwd(paste(folder,'plots',sep=''))

# extract mumax for plotting fit
pretty_spline_fits <-
  all_splines(value_mn ~ time | ID + biorep,
              data = tm_mtdt,
              spar = 0.5)
# extract mumax for table
all_spline_fits <-
  all_splines(value_mn ~ time | ID + biorep + full + strain + name + environment,
              data = tm_mtdt,
              spar = 0.5)

# plot spline fit
## change the par setting based on the amount of plots
jpeg(paste(
  format(Sys.time(), '%Y%m%d%H%M'),
  '_',
  project,
  '_',
  'plot_splines',
  '.jpeg',sep=''),
  width = 1188,
  height = 880)
par(mfrow = c(3, 6))
par(mar = c(2, 2, 1.5, 1.25))
plot(pretty_spline_fits, log='y')
dev.off()

# store results in dataframe
mumax<- results(all_spline_fits)
row.names(mumax)<- NULL

# remove blanks
mumax <- mumax[!mumax$strain=='blank',]
# remove mumax with rsquared below 0.95
mumax$mumax <- ifelse(mumax$r2 < 0.95, 0, mumax$mumax)

# calculate mean mumax
mumean<- ddply(mumax,c('ID','strain','name','environment'), 
               summarize,
               OD_mn=mean(y0),
               OD_sd=sd(y0),
               mumax_mn=mean(mumax),
               mumax_sd=sd(mumax))

# growthcurves - bars
p_mu <-
  ggplot(data = mumean, aes(x = ID, y = mumax_mn, fill = environment)) +
  ggtitle('Mean mumax of bioreps comparison') +
  xlab('Strain name (condition)') + 
  ylab('[mumax]')+
  geom_bar(stat='identity') +
  geom_errorbar(aes(ymin=mumax_mn-mumax_sd, ymax=mumax_mn+mumax_sd), width=.2,
                position=position_dodge(.9))+
  coord_flip() +
  theme(aspect.ratio=1) +
  theme(#legend.position = 'none',
    plot.title = element_text(face='bold', color='#000000', size=12, angle=0),
    axis.title.x = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.title.y = element_text(face='bold', color='#000000', size=10, angle=90),
    legend.title = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.text.y = element_text(face='italic', color='#000000', size=8, angle=0),
    axis.text.x = element_text(face='plain', color='#000000', size=8, angle=0),
    legend.text = element_text(face='italic', color='#000000', size=10, angle=0))

ggsave(p_mu,
       file = paste(
         format(Sys.time(), '%Y%m%d%H%M'),
         '_',
         project,
         '_',
         'plot_mumax',
         '.jpeg',
         sep = ''),
       width = 10, height = 8, units = "in", device='jpeg')
```

PLOT MUMAX (blank corrected)

For this plot we will first model the growth using a algorithm of Kahm et al. 2010 (see resources folder). The spline fitting is done with log-transformed data, assuming exponential growth at the time point of the maximum of its first derivative. The maximum growthrate is estimated based on this fit.

Actions:
Make sure to change the par setting in the first plot based on the amount of plots included.

```{r mumaxbc, include=FALSE, echo=TRUE}
# folder for plots
dir.create(paste(folder,'plots',sep=''))
setwd(paste(folder,'plots',sep=''))

# extract mumax for plotting fit
pretty_spline_fits <-
  all_splines(value_mn ~ time | ID + biorep,
              data = b_mtdt,
              spar = 0.5)
# extract mumax for table
all_spline_fits <-
  all_splines(value_mn ~ time | ID + biorep + full + strain + name + environment,
              data = b_mtdt,
              spar = 0.5)

# plot spline fit
## change the par setting based on the amount of plots
jpeg(paste(
  format(Sys.time(), '%Y%m%d%H%M'),
  '_',
  project,
  '_',
  'plot_splines_minb',
  '.jpeg',sep=''),
  width = 1188,
  height = 880)
par(mfrow = c(4, 4))
par(mar = c(2, 2, 1.5, 1.25))
plot(pretty_spline_fits, log='y')
dev.off()

# store results in dataframe
mumax<- results(all_spline_fits)
row.names(mumax)<- NULL

# remove blanks
mumax <- mumax[!mumax$strain=='blank',]
# remove mumax with rsquared below 0.95
mumax$mumax <- ifelse(mumax$r2 < 0.95, 0, mumax$mumax)

# calculate mean mumax
mumean<- ddply(mumax,c('ID','strain','name','environment'), 
               summarize,
               OD_mn=mean(y0),
               OD_sd=sd(y0),
               mumax_mn=mean(mumax),
               mumax_sd=sd(mumax))

# growthcurves - bars
p_mu <-
  ggplot(data = mumean, aes(x = ID, y = mumax_mn, fill = environment)) +
  ggtitle('Mean mumax of bioreps comparison') +
  xlab('Strain name (condition)') + 
  ylab('[mumax]')+
  geom_bar(stat='identity') +
  geom_errorbar(aes(ymin=mumax_mn-mumax_sd, ymax=mumax_mn+mumax_sd), width=.2,
                position=position_dodge(.9))+
  coord_flip() +
  theme(aspect.ratio=1) +
  theme(#legend.position = 'none',
    plot.title = element_text(face='bold', color='#000000', size=12, angle=0),
    axis.title.x = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.title.y = element_text(face='bold', color='#000000', size=10, angle=90),
    legend.title = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.text.y = element_text(face='italic', color='#000000', size=8, angle=0),
    axis.text.x = element_text(face='plain', color='#000000', size=8, angle=0),
    legend.text = element_text(face='italic', color='#000000', size=10, angle=0))

ggsave(p_mu,
       file = paste(
         format(Sys.time(), '%Y%m%d%H%M'),
         '_',
         project,
         '_',
         'plot_mumax_bcor',
         '.jpeg',
         sep = ''),
       width = 10, height = 8, units = "in", device='jpeg')
```

```{r requirements, include=TRUE, echo=TRUE}
# summary installed packages
sessionInfo()
# done
```